---
title: |
  | Applications of R for Finance
  | Course Project
author: |
  | Anthony Li
  | CID: 02308277
  | Business School, Imperial College London
date: "16-11-2022"
output: pdf_document
---

\newpage
\tableofcontents
\listoffigures
\listoftables
\newpage

In this assignment, we will load data from the file `compustat_sec_2019_2022.csv`,
which contains stock data for a large number of different companies.

I will be calculating and presenting the daily returns and the max monthly returns.
The September 2019 returns will be categorised into decile groups with the top 
decile group (top 10%) in September 2019 presented in a table.

The stock "NBCC (INDIA) LTD 01W" will be selected from the original data and a 
simple trading strategy will be applied to this stock. I will apply a Moving 
Average Crossover Strategy with the functions uploaded from a separate .R script. 
I will backtest this strategy 

## Load R packages

```{r Load libraries, message=FALSE, warning=FALSE}
library(knitr)
library(dplyr)
library(lubridate)
library(xts)
library(data.table)
library(dtplyr)
library(ggplot2)
library(PerformanceAnalytics)

# Importing another .R file which contains functions I will be using
source("rfun.R")
```

## Load and prepare data

We load and display the dataset as follows:

Load the entire dataset from `compustat_sec_2019_2022.csv`
```{r Load dataset}
# Loading data using fread which is much faster compared to read.csv hence should
# be used for large datasets. However, the result would be a table datatype rather
# than a data frame.
big_data <- fread("compustat_sec_2019_2022.csv", drop = c("sic", "exchg"))

# Displaying first 6 rows
head(big_data)
```

## Calculate daily returns
We use the lagged proceeds (prccd) in order to calculate daily returns of each stock 
in the dataset. I have defined the function to calculate returns in the rfun.R file.
```{r Daily Ret}
# I am applying pipe operators to modify the data. I have used the dtplyr so that
# I can apply the normal pipe operators to the data table datatype.
data <- big_data %>% 
  group_by(conm, iid) %>% 
  mutate(returns = simple_daily_return(prccd) ) %>% 
  ungroup() %>%
  as.data.table()

head(data)


```

## Calculate maximum daily returns for each month
We extract the month from each date using the floor_date() function and store it 
in a new column called return_month. We then look for the maximum daily returns 
for each month and store them in a new column called max_return. We only keep the 
rows where the maximum daily return is equal to the daily return. 

**Store the maximum daily returns in a new column called `max_return`**
```{r Maximum daily returns}
# Converting the datatype of datadate from Integer to Date
data$datadate <- as.Date(as.character(data$datadate), format = "%Y%m%d")

# Extract the month from each date using floor_date(), and store in a new column
data$return_month <- floor_date(data$datadate, unit = "month", 
                                week_start = getOption("lubridate.week.start",7))

# Create a new column named max_return with the maximum daily returns each month
max_daily_return <- data  %>%
  group_by(return_month, gvkey, iid, conm) %>%
  filter(is.finite(returns)) %>%
  mutate(max_return = max(returns, na.rm = TRUE) ) %>%
  ungroup() %>%
  as.data.table()
  
# Keep only the rows where the maximum daily return matches the daily return
max_daily_return.uni <- 
  max_daily_return[max_daily_return$returns == max_daily_return$max_return ,]

max_daily_return.uni1 <- max_daily_return.uni[!duplicated(max_daily_return.uni
[,c("return_month", "max_return", "iid", "conm", "gvkey")]), ]

```

```{r}
# Displaying max monthly returns for first 6 months
knitr::kable(
  head(max_daily_return.uni1), booktabs = TRUE,
  caption = 'Maximum monthly returns'
)
```


## Categorise into decile groups
Having obtained the maximum daily returns, we take a subset made up of the 
September 2019 maximum daily returns and we categorise them into decile groups 
using the cut() function.

**Categorise into deciles in a new column called `decile_cat`**
```{r Deciles}
# Extract a subset of the maximum daily returns (September 2019)
data_Sep2019 <- max_daily_return.uni1  %>% 
  filter(return_month=="2019-09-01") %>%
  as.data.table()


# Categorise into decile groups in a new column named decile_cat
data_Sep2019$decile_cat <- as.character(cut(data_Sep2019$returns,
                            quantile(data_Sep2019$returns, 
                            probs = seq(0,1, length = 11), na.rm = T, type = 5), 
                            include.lowest = TRUE, labels = c("10%", "20%", "30%"
                            ,"40%", "50%", "60%", "70%", "80%", "90%", "100%" )))

top_decile_group <- data_Sep2019[data_Sep2019$decile_cat == "100%",]
```

### Presenting the top 10% decile group (i.e. the 100% decile group)

```{r Choose stocks}
# Display in the final render
knitr::kable(
  head(top_decile_group), booktabs = TRUE,
  caption = 'Top 10% decile group in Sep 2019.'
)
```
## Extracting data for "NBCC (INDIA) LTD"

```{r Extracting Stock}
NBCC <- big_data %>%
  filter(conm == "NBCC (INDIA) LTD", iid == "01W") %>%
  as.data.frame() 

NBCC$datadate <- as.Date(as.character(NBCC$datadate), format = "%Y%m%d")

```



## Implementing Moving Average Crossover Strategy

```{r Trading Strategy}
NBCC.ts <- xts(NBCC$prccd, order.by = NBCC$datadate)
colnames(NBCC.ts) <- "NBCC"


# Using movingAvg function in the rfun.R script to calculate MA 
MA_100 <- movingAvg(NBCC.ts, 100)
MA_20 <- movingAvg(NBCC.ts, 20)

# Obtaining the a buy or sell signal depending on whether the short term MA has
# crossed the long term MA from above or below
signal <- sign(MA_20-MA_100)

# No moving average for first n observations where n is the long term MA length
signal[is.na(signal)] <- 0

# Plotting graph of the Moving Averages and the stock price
ggplot()+
  geom_line(NBCC, mapping = aes(x=datadate ,y=prccd, color="Stock Price"))+
  geom_line(NBCC, mapping = aes(x=datadate ,y=MA_100, color="Slow MA"))+
  geom_line(NBCC, mapping = aes(x=datadate ,y=MA_20, color="Fast MA"))+
  geom_line()+
  ggtitle("NBCC (INDIA) LTD 01W Share price") +
  ylab("Share Price") +
  xlab("Date") +
  scale_color_manual(name = c("Price", "Slow MA", "Fast MA"), 
      values = c("Stock Price" = "blue", "Slow MA" = "red", "Fast MA" = "green"))


results <- simpleBacktest(NBCC.ts, signal)

charts.PerformanceSummary(results)
```
## Analysis

The cumulative returns seem to be very volatile and occasionally reaches below 0.
However, the period ends with a positive cumulative return showing that the strategy
is profitable. 



